A load balancing scheme for massively multiplayer online games \thanks{This work was supported by the National Research Council (CNPq) and by the Coordination of Improvement of Higher Education (CAPES), both brazilian research funding agencies.}

### 2009-04-10 textos/dissertation/MTAP_LaTeX/pub/tex/latex/svjour3/global/balancing.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{algorithmic}
\usepackage{algorithm}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}
%[TODO:ajustar TODAS as \refs]
\title{A load balancing scheme for massively multiplayer onlne games \thanks{This work was supported by the National Research Council (CNPq) and by the Coordination of Improvement of Higher Education (CAPES), both brazilian research funding agencies.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\begin{abstract}
MMOGs are applications which require much network bandwidth to function properly. In a distributed MMOG server architecture, with heterogeneous resources, the server nodes may become easily overloaded by the high demand from the players for state updates. In this work, we propose a balancing scheme which has three main goals: cope with the most important load on the servers, which is the network traffic, allocate load on the servers proportionally to the each one's power, and reduce as much as possible the overhead from the distribution. It is is divided in 3 phases: local selection of servers, balancing and refinement. Four algorithms were proposed, from which \ggp\ is the best for overhead reduction and \ggpf\ is the most suited for reducing player migrations between servers.

@@

The main characteristic of massively multiplayer games is the large number of players, having dozens, or even hundreds, of thousands participants simultaneously \cite{schiele2007rpp}. This large number of players interacting with one another generates a traffic on the support network which grows quadratically compared to the number of players, in the worst case.

When using a client-server architecture, it is necessary that the server intermediates the communication between each pair of players -- assuming that the game is intended to provide guarantees of consistency and resistance to cheating. Obviously, this server will have a large communication load, thus, it must have enough resources (available bandwidth) to meet the demand of the game. We consider here that the main resource to analyse is the available bandwidht, for this is the current bottleneck for MMOGs \cite{feng2007wnn}.

The question is, when you use a distributed server, you must delegate to each server node a load proportional to its power. Thus, no matter to which server each player is connected, their game experience will be similar, regarding the response time for their actions and the time it takes to be notified of actions from other players as well as of state changes in the virtual environment of the game.

An initial idea might be to distribute the players between servers, so that the number of players on each server would be proportional to that server's bandwidth. However, this distribution would not work, for the burden caused by players also depends on how they are interacting with one another. For example, if the avatars of two players are too distant from each other, probably there will be no interaction between them and therefore the server needs only to update each one of them with the result of their own actions. However, if these avatars are close to each other, each player should be updated not only of his own actions, but also of the actions of the other player.

@@

\begin{algorithm}
%\renewcommand{\ALG}{Algoritmo}
\caption{\ggpk}
\label{alg:ggpk}
\begin{algorithmic}[1]
	 \STATE $\wtodiv \leftarrow 0$
	 \STATE $\freecap \leftarrow 0$
	 \FOR{each $R$ in $\rlist$}
	 		\STATE $\wtodiv \leftarrow \wtodiv + w_r(R)$
	 		\STATE $\freecap \leftarrow \freecap + p(s(R))$
	 		\STATE $c \leftarrow $ heaviest cell from $R$ \label{alg:ggpk:saveheaviestcell}
	 		\STATE temporarily free all cells from $R$
	 		\STATE $R \leftarrow R \cup \{c\}$ \label{alg:ggpk:getbackheaviestcell}
	 \ENDFOR
	 \STATE sort $\rlist$ in decreasing $p(s(R))$ order \label{alg:ggp:sortregions}
	 \FOR{each region $R$ in $\rlist$}
	 		\STATE $\wshare \leftarrow \wtodiv \times \frac{p(s(R))}{\freecap}$ \label{alg:ggp:fraction}
	 		\WHILE{$w_r(R) < \wshare$} \label{alg:ggp:whilefraction}
	 		   \IF{there is any cell from $R$ neighboring a free cell}
	 		   		\STATE $R \leftarrow R \cup \{$neighbor free cell with the highest $Int_c(C)$$\}$ \label{alg:ggp:freeneighbor}
	 		   \ELSIF{there is any free cell}
	 		   		\STATE $R \leftarrow R \cup \{$heaviest free cell$\}$ \label{alg:ggp:freeheavycell}
	 		   \ELSE
	 		   		\STATE stop. no more free cells.
	 		   \ENDIF
	 		\ENDWHILE
	 \ENDFOR
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\ggpf}
\label{alg:ggpf}
\begin{algorithmic}[1]
	 \STATE $\wtodiv \leftarrow 0$
	 \STATE $\freecap \leftarrow 0$
	 \FOR{each region $R$ in $\rlist$}
	 		\STATE $\wtodiv \leftarrow \wtodiv + w_r(R)$
	 		\STATE $\freecap \leftarrow \freecap + p(s(R))$
	 		\STATE $\clist \leftarrow $ list of cells from $R$ in increasing order of weight
	 		\WHILE{$frac_r(R) > frac_p(s(R))$}
	 			 \STATE $C \leftarrow$ first element from $\clist$
	 			 \STATE remove $C$ from $R$
	 			 \STATE remove $C$ from $\clist$
	 		\ENDWHILE
	 \ENDFOR
	 \STATE sort $\rlist$ in increasing order of $u(s(R))$ \label{alg:ggpf:usageorder}
	 \FOR{each region $R$ in $\rlist$}
	 		\STATE $\wshare \leftarrow \wtodiv \times \frac{p(s(R))}{\freecap}$
	 		\WHILE{$w_r(R) < \wshare$} \label{alg:ggp:whilefraction}
 		   	 \IF{there is any cell from $R$ neighboring a free cell}
	 		   		\STATE $R \leftarrow R \cup \{$neighbor free cell with the highest $Int_c(C)$$\}$ \label{alg:ggp:freeneighbor}
	 		   \ELSIF{there is any free cell}
	 		   		\STATE $R \leftarrow R \cup \{$heaviest free cell$\}$ \label{alg:ggp:freeheavycell}
	 		   \ELSE
	 		   		\STATE stop. no more free cells.
	 		   \ENDIF
	 		\ENDWHILE
	 \ENDFOR\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/baloverall}
  \caption{Overall comparison of proposed load balancing algorithms}
  \label{gra:baloverall}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/baloverhead}
  \caption{Overhead introduced by each balancing algorithm during the game session}
  \label{gra:baloverhead}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/balmigxtb}
  \caption{Players migrating while still or walking depending on algorithm}
  \label{gra:balmigxtb}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/baldeviation}
  \caption{Deviation of the ideal usage value}
  \label{gra:baldeviation}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{algorithmic}
\usepackage{algorithm}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}
%[TODO:ajustar TODAS as \refs]
\title{A load balancing scheme for massively multiplayer online games \thanks{This work was supported by the National Research Council (CNPq) and by the Coordination of Improvement of Higher Education (CAPES), both brazilian research funding agencies.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\begin{abstract}
MMOGs are applications which require much network bandwidth to function properly. In a distributed MMOG server architecture, with heterogeneous resources, the server nodes may become easily overloaded by the high demand from the players for state updates. Much work has been done in order to distribute the load on the server nodes, but this load is usually defined as the number of players on each server, what is not an ideal measure. Also, the heterogeneity of the system is frequently overlooked. In this work, we propose a balancing scheme which has two main goals: allocate load on the servers proportionally to the each one's power and reduce as much as possible the overhead from the distribution, considering the load as the occupied bandwidth of each server. It is is divided in three phases: local selection of servers, balancing and refinement. Four algorithms were proposed, from which \ggp\ is the best for overhead reduction and \ggpf\ is the most suited for reducing player migrations between servers. We also make a review of related works and some comparisons were made.

@@

The main characteristic of massively multiplayer games is the large number of players, having dozens, or even hundreds, of thousands of participants simultaneously \cite{schiele2007rpp}. This large number of players interacting with one another generates a traffic on the support network which may grow quadratically compared to the number of players, in the worst case (Figure \ref{fig:quadratic}).

\begin{figure}
  \centering
  \includegraphics[width=1.0\linewidth]{images/avatarsdistribution}
  \caption{Distribution of avatars with and without hotspots}
  \label{fig:quadratic}
\end{figure}

When using a client-server architecture, it is necessary that the server intermediates the communication between each pair of players -- assuming that the game is intended to provide guarantees of consistency and resistance to cheating. Obviously, this server will have a large communication load, thus, it must have enough resources (available bandwidth) to meet the demand of the game. We consider here that the main resource to analyse is the available bandwidth, for this is the current bottleneck for MMOGs \cite{feng2007wnn}.

The problem is that, when using a distributed server, it must be delegated to each server node a load proportional to its power. Thus, no matter to which server each player is connected, their game experience will be similar, regarding the response time for their actions and the time it takes to be notified of actions from other players as well as of state changes in the virtual environment of the game.

An initial idea might be to distribute the players among servers, so that the number of players on each server would be proportional to that server's bandwidth. However, this distribution would not work, for the burden caused by players also depends on how they are interacting with one another. For example, if the avatars of two players are too distant from each other, probably there will be no interaction between them and therefore the server needs only to update each one of them with the result of their own actions. However, if these avatars are close to each other, each player should be updated not only of his own actions, but also of the actions of the other player.

@@

\begin{algorithm}
%\renewcommand{\ALG}{Algoritmo}
\caption{\ggpk}
\label{alg:ggpk}
\begin{algorithmic}[1]
	 \STATE $\wtodiv \leftarrow 0$
	 \STATE $\freecap \leftarrow 0$
	 \FOR{each $R$ in $\rlist$}
	 		\STATE $\wtodiv \leftarrow \wtodiv + w_r(R)$
	 		\STATE $\freecap \leftarrow \freecap + p(s(R))$
	 		\STATE $c \leftarrow $ heaviest cell from $R$ \label{alg:ggpk:saveheaviestcell}
	 		\STATE temporarily free all cells from $R$
	 		\STATE $R \leftarrow R \cup \{c\}$ \label{alg:ggpk:getbackheaviestcell}
	 \ENDFOR
	 \STATE sort $\rlist$ in decreasing $p(s(R))$ order \label{alg:ggp:sortregions}
	 \FOR{each region $R$ in $\rlist$}
	 		\STATE $\wshare \leftarrow \wtodiv \times \frac{p(s(R))}{\freecap}$
	 		\WHILE{$w_r(R) < \wshare$} \label{alg:ggpk:whilefraction}
	 		   \IF{there is any cell from $R$ neighboring a free cell}
	 		   		\STATE $R \leftarrow R \cup \{$neighbor free cell with the highest $Int_c(C)$$\}$
	 		   \ELSIF{there is any free cell}
	 		   		\STATE $R \leftarrow R \cup \{$heaviest free cell$\}$ \label{alg:ggpk:freeheavycell}
	 		   \ELSE
	 		   		\STATE stop. no more free cells.
	 		   \ENDIF
	 		\ENDWHILE
	 \ENDFOR
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\ggpf}
\label{alg:ggpf}
\begin{algorithmic}[1]
	 \STATE $\wtodiv \leftarrow 0$
	 \STATE $\freecap \leftarrow 0$
	 \FOR{each region $R$ in $\rlist$}
	 		\STATE $\wtodiv \leftarrow \wtodiv + w_r(R)$
	 		\STATE $\freecap \leftarrow \freecap + p(s(R))$
	 		\STATE $\clist \leftarrow $ list of cells from $R$ in increasing order of weight
	 		\WHILE{$frac_r(R) > frac_p(s(R))$}
	 			 \STATE $C \leftarrow$ first element from $\clist$
	 			 \STATE remove $C$ from $R$
	 			 \STATE remove $C$ from $\clist$
	 		\ENDWHILE
	 \ENDFOR
	 \STATE sort $\rlist$ in increasing order of $u(s(R))$ \label{alg:ggpf:usageorder}
	 \FOR{each region $R$ in $\rlist$}
	 		\STATE $\wshare \leftarrow \wtodiv \times \frac{p(s(R))}{\freecap}$
	 		\WHILE{$w_r(R) < \wshare$} \label{alg:ggpf:whilefraction}
 		   	 \IF{there is any cell from $R$ neighboring a free cell}
	 		   		\STATE $R \leftarrow R \cup \{$neighbor free cell with the highest $Int_c(C)$$\}$
	 		   \ELSIF{there is any free cell}
	 		   		\STATE $R \leftarrow R \cup \{$heaviest free cell$\}$ \label{alg:ggpf:freeheavycell}
	 		   \ELSE
	 		   		\STATE stop. no more free cells.
	 		   \ENDIF
	 		\ENDWHILE
	 \ENDFOR\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=1.0\linewidth]{images/baloverall_ah}
  \caption{Overall comparison of proposed load balancing algorithms}
  \label{gra:baloverall}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/baloverhead_ah}
  \caption{Overhead introduced by each balancing algorithm during the game session}
  \label{gra:baloverhead}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/balmigxtb_ah}
  \caption{Players migrating while still or walking depending on algorithm}
  \label{gra:balmigxtb}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/baldeviation_ah}
  \caption{Deviation of the ideal usage value}
  \label{gra:baldeviation}
\end{figure}
